Hierarchy Builder

### 2020-04-24 coqws/easychair.tex
cleanup

\section{Context}

There is no questioning the importance of abstract algebra in the mathematical
vernacular. And it is with no surprise that libraries of machine checked code
are, nowadays, organized around hierarchies of structures. Unfortunately the
language of Type Theory and the features provided by the Coq system make the
construction of a hierarchy hard even for expert users. The difficulty
begins with the non orthogonal choices of the encoding style, from bundled to
unbundled, and inference mechanism, from Type Classes to Canonical Structures.
To this one may add the concerns about performance and finally the ones about
the usability of the final hierarchy.

In this talk we give an extensive demo of the Hierarchy Builder (\HB{})
extension for the Coq system version 8.10 and above.
\HB{} gives the library designer a language to describe the building blocks
of algebraic structures and to assemble them into a hierarchy. Similarly it
provides the final user linguistic constructs to build instances (examples) of
structures and to teach the elaborator of Coq how to take advantage of this
knowledge during type inference. Finally \HB{} lets the library designer
improve the usability of his library by providing alternative interfaces
to the primitive ones, a feature that can also be used to accommodate changes
without breaking user code.

The details of \HB{}, including its compilation to the (Flat) Packed Classes
discipline and its implementation in the Elpi extension language for Coq,
are given in~\cite{cohen:hal-02478907}.
In this talk we focus on how to take advantage
of \HB{} in a Coq development, and we plan to not detail how it works
behind the scenes, unless specific questions arise from the audience.

@@

For space constraints we omit the 30 lines in which we use
the \coq{HB.factory} command to declare an alternative interface
to build examples of the \coq{Ring} structure (the green
arrow in the picture) by providing all proofs at once but for the
one on the commutativity of the Monoid operation:
the other Ring axioms imply it~\cite{nearrings}. Factories can also be used to
preserve backward compatibility when an hierarchy evolves to accommodate
new intermediate  structures.

>>>>

\section{Context and Talk proposal}

There is no questioning the importance of abstract algebra in the mathematical
vernacular. And it is with no surprise that libraries of machine checked code
are, nowadays, organized around hierarchies of structures. Unfortunately the
language of Type Theory and the features provided by the Coq system make the
construction of a hierarchy hard even for expert users. The difficulty
begins with the non-orthogonal choices of the encoding style, from bundled to
unbundled, and inference mechanism, from Type Classes to Canonical Structures.
To this, one may add the concerns about performance and about
the usability, by a non expert, of the final hierarchy.

In this talk we give an extensive demo of the Hierarchy Builder (\HB{})
extension for the Coq system version 8.10 and above.
\HB{} gives the library designer a language to describe the building blocks
of algebraic structures and to assemble them into a hierarchy. Similarly it
provides the final user linguistic constructs to build instances (examples) of
structures and to teach the elaborator of Coq how to take advantage of this
knowledge during type inference. Finally \HB{} lets the library designer
improve the usability of his library by providing alternative interfaces
to the primitive ones, a feature that can also be used to accommodate changes
to the hierarchy without breaking user code.

The details of \HB{}, including its compilation to the (Flat) Packed Classes
discipline and its implementation in the Elpi extension language for Coq,
are given in~\cite{cohen:hal-02478907}.
In this talk we focus on how to use \HB{} in a Coq development, and we plan
to not detail how it works behind the scenes, unless specific questions
arise from the audience.

@@

For space constraints we omit the declaration of the green arrow
in the picture (about 30 lines) which is an alternative interface
to build examples of the \coq{Ring} structure by providing all
proofs at once. In that case the user is spared to prove the commutativity
of the Monoid operation since it is implied by the other axioms~\cite{nearrings}.
